---
title: "Data Tidying"
author: "Larissa Lee"
date: "5/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries}
library(dplyr)
library(tidyr)
```

Read in data from the [KNB](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1)

```{r get_data}
catch_df <-read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
```

Selected columns that I wanted, moved from wide to long format, and renamed a column

```{r Remove Summary & Notes}
catch_df<- catch_df %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>%
  gather(key = Species, value = Catch, -Region, -Year) %>%
  rename(catch_thousands = Catch)
```

Found a typo in original data, corrected it, converted 'catch_thousands' column to numeric and created a new column with units # of fish. used code below:

{r FindError}
test <-as.integer(catch_df$catch_thousands)
which(is.na(test == TRUE))
catch_df[401, ]


```{r EditError_AddCatchColumn}
catch_df <- catch_df %>%
    mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>%
  mutate(catch_thousands = as.integer(catch_thousands)) %>%
  mutate(catch = catch_thousands * 1000)
head(catch_df, 30)
```

# Analysis

```{r Group}
catch_df %>% 
  group_by(Species, Year) %>%
  summarize(mean_catch = mean(catch), num_obs = n())
```

```{r Filter}
catch_df %>%
  filter(Species == "Chinook") %>% # look at only Chinook
  group_by(Region) %>% #group by region 
  summarise(mean_catch = mean(catch)) %>% #summarize by catch
  arrange(desc(mean_catch)) #arrange in descending order
```
# Joins

```{r Joins}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```

```{r}
region_defs <- region_defs %>%
  select(code, mgmtArea) %>%
  rename(Region = code, Region_Name = mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_df, region_defs)
head(catch_joined)
```

